---
title: "Snowpack in the Western U.S."
output:
  flexdashboard::flex_dashboard
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load-packages}
library(tidyverse)
library(flexdashboard)
library(crosstalk)
library(leaflet)
```

```{r apr-extract-transform}

apr_all <- read_csv("snowpack-april.csv")

apr_updt = apr_all %>%
  mutate(change = ifelse(trend >= 0,"Positive","Negative"))

apr_shared = SharedData$new(apr_updt)
```

```{r timing-extract-transform}

timing_all <- read_csv("snowpack-timing.csv")

timing_updt = timing_all %>%
  filter(state != "AK") %>%
  mutate(change = ifelse(timing >= 0,"Positive","Negative"))

timing_shared = SharedData$new(timing_updt)
```

Amount
=====================================

This map shows how much April snowpack has changed from 1982 to 2020.

Each circle represents a different site in the western United States.

- Negative percentages, shown in red, indicate decreased snowpack.
- Positive percentages, shown in blue, indicate increased snowpack.

Column {data-width=25}
-------------------------------------
### Filters

```{r apr-filter}
filter_checkbox(
  id = "change",
  label = "Change",
  sharedData = apr_shared,
  group = ~change
)
```

Column {data-width=100}
-------------------------------------
### Interactive Map
```{r apr-map}
apr_map <- leaflet(data = apr_shared)

pal <- colorNumeric(
  palette = "RdBu",
  domain = c(-110,110))

apr_map %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = 5,
    stroke = TRUE,
    color = "black",
    weight = 1,
    fillColor = ~pal(trend),
    fillOpacity = 0.75,
    popup = paste0("<b>Latitude: </b>", format(round(apr_shared$data()$latitude, digits=2), nsmall=2), "<br>",
                   "<b>Longitude: </b>", format(round(apr_shared$data()$longitude, digits=2), nsmall=2), "<br>",
                   "<b>Amount Changed: </b>", format(round(apr_shared$data()$trend, digits=2), nsmal=2), "%")
  ) %>%
  addProviderTiles(providers$Stamen.Terrain) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~trend,
    title = "Amount Changed (%)",
    opacity = 1
  )
```
Timing
=====================================

This map shows how the date when snowpack reaches its maximum amount has changed from 1982 to 2020.

Each circle represents a different site in the western United States.

- Negative values, shown in red, indicate a shift to earlier days.
- Positive values, shown in blue, indicate a shift to later days.

Column {data-width=25}
-------------------------------------
### Filters

```{r timing-filter}
filter_checkbox(
  id = "change",
  label = "Change",
  sharedData = timing_shared,
  group = ~change
)
```

Column {data-width=100}
-------------------------------------
### Interactive Map
```{r timing-map}
timing_map <- leaflet(data = timing_shared)

pal <- colorNumeric(
  palette = "RdBu",
  domain = c(-42,42))

timing_map %>%
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = 5,
    stroke = TRUE,
    color = "black",
    weight = 1,
    fillColor = ~pal(timing),
    fillOpacity = 0.75,
    popup = paste0("<b>Latitude: </b>", format(round(timing_shared$data()$latitude, digits=2), nsmall=2), "<br>",
                   "<b>Longitude: </b>", format(round(timing_shared$data()$longitude, digits=2), nsmall=2), "<br>",
                   "<b>Timing: </b>", format(round(timing_shared$data()$timing, digits=2), nsmal=2), " days")
  ) %>%
  addProviderTiles(providers$Stamen.Terrain) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = ~timing,
    title = "Timing Shift (Days)",
    opacity = 1
  )
```

Notes
=====================================

By [Jack J. Lee](https://www.jackjleescience.com)

Data from the [U.S. Environmental Protection Agency](https://www.epa.gov/climate-indicators/climate-change-indicators-snowpack)

Files available on [Github](https://github.com/jackjlee/snowpack).
